Artin Group Presentations Arising from Cluster Algebras

### 2014-08-17 Paper.tex
Updated changes for sections 1 and 2

In \cite{FZ02}, Fomin and Zelevinsky first introduced the concept of cluster algebras. Barot and Marsh extended Fomin and Zelevinsky's results in \cite{BM13}, providing a presentation of the reflection group associated to a Dynkin diagram with generators that correspond to elements of a companion basis associated to a seed of a finite type cluster algebra. They define generator relations corresponding to chordless cycles arising in diagrams of finite type in order to give a Coxeter group presentation for these diagrams. They also proved that this Coxeter group presentation is invariant up to isomorphism under the mutation equivalence relation.That is, given a diagram $\Gamma$ and a diagram mutation equivalent to $\Gamma$, denoted $\Gamma^{\prime} = \mu_k(\Gamma)$, they proved that $W_{\Gamma} \cong W_{\Gamma^{\prime}}$, where $W_{\Gamma}$ and $W_{\Gamma^{\prime}}$ are the group representations corresponding to $\Gamma$ and $\Gamma^{\prime}$, respectively.

@@

In our paper, we define $A_{\Gamma}$ to be the Artin group presentation arising from a cluster algebra, where $\Gamma$ is the diagram associated to the cluster algebra. We provide the necessary relations for the generators of the group, and show that these relations hold under mutations of chordless cycles in a diagram. Our main result is to show that this Artin group presentation is invariant up to isomorphism under the mutation equivalence relation. We state the result here, but present the detailed proof in Section \ref{sec:main_result}.

@@

We begin by introducing some preliminary notations and definitions which will aid the reader in understanding the results in the following sections. For further references on cluster algebras, we refer the reader to \cite{FZ02} and \cite{FZ03} and for a more detailed description of Artin group presentations, we direct attention to \cite{FN61}. We also provide references to several lemmas and propositions from \cite{BM13} which were helpful in formulating our own results. \\

The initial introduction of cluster algebras by Fomin and Zelevinsky was aimed at making further strides in the areas of representation theory, Lie theory, and total positivity. Since then, the study of cluster algebras has provided a motivation for applications in various other areas of mathematics, including quiver representations. Of particular interest were \textit{finite type} cluster algebras. In the sequel to their introductory paper (\cite{FZ03}), Fomin and Zelevinsky introduce the concept of \textit{mutation equivalence} between diagrams, proving that a connected graph is mutation equivalent to an oriented Dynkin diagram if and only if all mutation equivalent graphs have edge weights not exceeding 3. In particular, this proves that finite type cluster algebras can be classified by Dynkin diagrams. \\
A \textit{cluster algebra} is an integral domain which can be generated by a set of elements called \textit{cluster variables} that satisfy certain exchange relations. Following the style of \cite{FZ02} and \cite{BM13}, we will define cluster algebras in terms of \textit{skew-symmetrisable} matrices (that is, a matrix $B$ such that there exists a diagonal matrix $D$ of the same size with $D_{ii} >0$ such that $DB$ is skew-symmetric). Let $\mathbb{F} = \mathbb{Q}(u_1, u_2, \ldots, u_n)$ be the field of rational functions in $n$ indeterminates over $\mathbb{Q}$. We will define an \textit{intial seed} for the cluster algebra to be a fixed pair $(\textbf{x}, B)$, where $\textbf{x} = \{x_1, \ldots, x_n\}$ is a free generating set of $\mathbb{F}$ and $B$ is an $n \times n$ skew-symmetric matrix. Define $x_k^{\prime} \in \mathbb{F}$ by the \textit{exchange relation}
\begin{displaymath}
x_k^{\prime}x_k = \prod_{B_{ik} > 0}{x_i^{B_{ik}}} + \prod_{B_{ik} < 0}{x_i^{-B_{ik}}}
\end{displaymath}
Then, given an initial seed $(\textbf{x}, B)$ and $k \in {1,2,\ldots,n}$, we can define a \textit{mutation} of the seed at $k$, denoted $\mu_k(\boldsymbol{x}, B) = (\textbf{x}^{\prime}, B^{\prime})$ where:
\begin{displaymath}
B_{ij}^{\prime} = \begin{cases} - B_{ij} & \mbox{ if } i = k \mbox{ or } j = k;\\
B_{ij} + \frac{|B_{ik}|B_{kj} + B_{ik}|B_{kj}|}{2} & \mbox{ otherwise. }\\
\end{cases}
\end{displaymath}

and $\textbf{x}^{\prime} = \left\{ x_1, x_2, \ldots, x_{k-1}, x_k^{\prime}, x_{k+1}, \ldots, x_n \right\}$.
Such a mutation or a sequence of such mutations generate \textit{seeds} which in turn generate all cluster variables in that, for each $\textbf{x} = \left\{ x_1, \ldots, x_n \right\}$ corresponding to a seed of the cluster algebra, the entries $x_i$ are the cluster variables. \\

 A cluster algebra is said to be of \textit{finite type} if the number of cluster variables that generate it is finite (if it has finitely many seeds). For each finite type cluster algebra, we can associate to its corresponding skew-symmetrisable matrix an edge-weighted, oriented graph, called a \textit{diagram}. We will often denote this diagram by $\Gamma$, and the vertex set of $\Gamma$ by $\V$. We will denote two connected vertices by $i \rightarrow j$, or by $i\--\ j$ if the orientation is not specified. The diagram is determined by, for $i, j \in \V$, $i \xrightarrow{w} j$ if and only if $B_{ij} > 0$ and $w = |B_{ij}B_{ji}|$ is the weight of the edge. A skew-symmetrisable matrix $B$ is \textit{2-finite} if $|B_{ij}B_{ji}| \leq 3$ for $i, j \in \left\{ 1, \ldots, n \right\}$. By \cite[7.5]{FZ02}, we have that if $B$ is 2-finite, all 3-cycles in the unoriented graph underlying our diagram must be oriented cyclically. \\

@@

Then, they define $W(\Gamma)$ to be the group generated by $s_i$, for $i \in \V$, under the following relations. Note that $e$ will denote the identity element of $W(\Gamma)$.
\begin{enumerate}
\item $s_i^2 = e$ for all $i$;
\item $(s_is_j)^{m_{ij}} = e$ for all $i \neq j$;
\item For any chordless cycle (as defined in Definition \ref{chordlesscycle} $C$ in $\Gamma$, where
\begin{displaymath}
C = i_0 \xrightarrow{w_1} i_1 \xrightarrow{w_2} \cdots \xrightarrow{w_{d-1}} i_{d-1} \xrightarrow{w_0} i_0
\end{displaymath}
and all of the weights $w_k$ are $1$ or $w_0 = 2$, we have
\begin{displaymath}
(s_{i_0}s_{i_1}\cdots s_{i_{d-2}}s_{i_{d-1}}s_{i_{d-2}}\cdots s_{i_1})^2 =e.
\end{displaymath}
%\todo{Maybe it's worth presenting definition 2.4 before this, and then saying it is equivalent to removing certain the (R3)(b) relations. Don't they actually present the definition you refer to in Definition 2.4 as the presentation? You say it is an alteration there, but I think it's the original definition}
\end{enumerate}

@@

In Section 3 of \cite{BM13}, Barot and Marsh provide an alteration of the group $W(\Gamma)$ in order to extend the group definition to any diagram of finite type. The group they define is as follows:

@@

The rest of the paper will be devoted to building up analogous relations, defined in ~\ref{grp def} to prove a similar result in the case of Artin groups. For $\Gamma$ a diagram of finite type, we define the Artin group associated to $\Gamma$ as in Section \ref{sec:defn_artingroup}. We will then use the group relations presented in this definition to prove the mutation invariance of $A_{\Gamma}$ in Section \ref{sec:main_result}.

>>>>

In \cite{FZ02}, Fomin and Zelevinsky first introduced the concept of cluster algebras. Barot and Marsh extended Fomin and Zelevinsky's results in \cite{BM13}, providing a presentation of the reflection group associated to a Dynkin diagram with generators that correspond to elements of a companion basis associated to a seed of a finite type cluster algebra. They define generator relations corresponding to chordless cycles arising in diagrams of finite type in order to give a Coxeter group presentation for these diagrams. They also proved that this Coxeter group presentation is invariant up to isomorphism under the mutation equivalence relation. That is, given a diagram $\Gamma$ and a diagram mutation equivalent to $\Gamma$, denoted $\Gamma^{\prime} = \mu_k(\Gamma)$, they proved that $W_{\Gamma} \cong W_{\Gamma^{\prime}}$, where $W_{\Gamma}$ and $W_{\Gamma^{\prime}}$ are the group presentations corresponding to $\Gamma$ and $\Gamma^{\prime}$, respectively.

@@

In our paper, we define $A_{\Gamma}$ to be the Artin group presentation arising from a cluster algebra, where $\Gamma$ is the diagram associated to the cluster algebra. We provide the necessary relations for the generators of the group, and show that these relations hold under mutations of vertices in a diagram. Our main result is to show that this Artin group presentation is invariant, up to isomorphism, under the mutation equivalence relation. We state the result here, but present the detailed proof in Section \ref{sec:main_result}.

@@

We begin by introducing some preliminary notations and definitions which will aid the reader in understanding the results in the following sections. For further references on cluster algebras, we refer the reader to \cite{FZ02, FZ03}, and for a more detailed description of Artin group presentations, we direct attention to \cite{FN61}. We also provide references to several lemmas and propositions from \cite{BM13} which were helpful in formulating our own results. \\

The initial introduction of cluster algebras by Fomin and Zelevinsky was aimed at making further strides in the areas of representation theory, Lie theory, and total positivity. Since then, the study of cluster algebras has provided a motivation for applications in various other areas of mathematics, including quiver representations. Of particular interest were \textit{finite type} cluster algebras, those with a finite number of distinguished generators. In the sequel to their introductory paper (\cite{FZ03}), Fomin and Zelevinsky introduce the concept of \textit{mutation equivalence} between diagrams, proving that a connected graph is mutation equivalent to an oriented Dynkin diagram if and only if all mutation equivalent graphs have edge weights not exceeding 3. In particular, this proves that finite type cluster algebras can be classified by Dynkin diagrams. \\

A \textit{cluster algebra} is an integral domain which can be generated by a set of elements called \textit{cluster variables} that satisfy certain exchange relations. Following the style of \cite{FZ02} and \cite{BM13}, we will define cluster algebras in terms of \textit{skew-symmetrisable} matrices (that is, a matrix $B$ such that there exists a diagonal matrix $D$ of the same size with $D_{ii} >0$ such that $DB$ is skew-symmetric). Let $\mathbb{F} = \mathbb{Q}(u_1, u_2, \ldots, u_n)$ be the field of rational functions in $n$ indeterminates over $\mathbb{Q}$. We will define an \textit{intial seed} for the cluster algebra to be a fixed pair $(\textbf{x}, B)$, where $\textbf{x} = \{x_1, \ldots, x_n\}$ is a free generating set of $\mathbb{F}$ and $B$ is an $n \times n$ skew-symmetric matrix. Define $x_k^{\prime} \in \mathbb{F}$ by the \textit{exchange relation}
\begin{displaymath}
x_k^{\prime}x_k = \prod_{B_{ik} > 0}{x_i^{B_{ik}}} + \prod_{B_{ik} < 0}{x_i^{-B_{ik}}}
\end{displaymath}
Then, given an initial seed $(\textbf{x}, B)$ and $k \in {1,2,\ldots,n}$, we can define a \textit{mutation} of the seed at $k$, denoted $\mu_k(\boldsymbol{x}, B) = (\textbf{x}^{\prime}, B^{\prime})$ where:
\begin{displaymath}
B_{ij}^{\prime} = \begin{cases} - B_{ij} & \mbox{ if } i = k \mbox{ or } j = k;\\
B_{ij} + \frac{|B_{ik}|B_{kj} + B_{ik}|B_{kj}|}{2} & \mbox{ otherwise. }\\
\end{cases}
\end{displaymath}
and $\textbf{x}^{\prime} = \left\{ x_1, x_2, \ldots, x_{k-1}, x_k^{\prime}, x_{k+1}, \ldots, x_n \right\}$.
Such a mutation or a sequence of such mutations generate \textit{seeds} which in turn generate all cluster variables in that, for each $\textbf{x} = \left\{ x_1, \ldots, x_n \right\}$ corresponding to a seed of the cluster algebra, the entries $x_i$ are the cluster variables. \\

 A cluster algebra is said to be of \textit{finite type} if the number of cluster variables that generate it is finite (if it has finitely many seeds). For each finite type cluster algebra, we can associate to its corresponding skew-symmetrisable matrix an edge-weighted, oriented graph, called a \textit{diagram}. We will often denote this diagram by $\Gamma$, and the vertex set of $\Gamma$ by $\V$. We will denote two connected vertices by $i \rightarrow j$, or by $i\--\ j$ if the orientation is not specified. The diagram is determined by the following: for $i, j \in \V$, $i \xrightarrow{w} j$ if and only if $B_{ij} > 0$ and $w = |B_{ij}B_{ji}|$ is the weight of the edge. A skew-symmetrisable matrix $B$ is \textit{2-finite} if $|B_{ij}B_{ji}| \leq 3$ for $i, j \in \left\{ 1, \ldots, n \right\}$. By \cite[7.5]{FZ02}, we have that if $B$ is 2-finite, all 3-cycles in the unoriented graph underlying our diagram must be oriented cyclically. \\

@@

Then, they define $W(\Gamma)$ to be the group generated by $s_i$, for $i \in \V$, under the following relations. Note that $e$ will denote the identity element of $W(\Gamma)$.
\begin{enumerate}
\item $s_i^2 = e$ for all $i$;
\item $(s_is_j)^{m_{ij}} = e$ for all $i \neq j$;
\item For any chordless cycle (as defined in Definition \ref{chordlesscycle}) $C$ in $\Gamma$, where
\begin{displaymath}
C = i_0 \xrightarrow{w_1} i_1 \xrightarrow{w_2} \cdots \xrightarrow{w_{d-1}} i_{d-1} \xrightarrow{w_0} i_0
\end{displaymath}
and all of the weights $w_k$ are $1$ or $w_0 = 2$, we have
\begin{displaymath}
(s_{i_0}s_{i_1}\cdots s_{i_{d-2}}s_{i_{d-1}}s_{i_{d-2}}\cdots s_{i_1})^2 =e.
\end{displaymath}
%\todo{Maybe it's worth presenting definition 2.4 before this, and then saying it is equivalent to removing certain the (R3)(b) relations. Don't they actually present the definition you refer to in Definition 2.4 as the presentation? You say it is an alteration there, but I think it's the original definition}
\end{enumerate}

@@

In Section 3 of \cite{BM13}, Barot and Marsh provide an alteration of the group $W(\Gamma)$ in order to extend the group definition to any diagram of finite type. More specifically, they provide relations such that $W_{\Gamma}$, as defined below, is isomorphic to $W(\Gamma)$ (\cite[Proposition 4.5]{BM13}).

@@

The rest of the paper will be devoted to building up analogous relations, defined in Definition~\ref{grp def} to prove a similar result in the case of Artin groups. For $\Gamma$ a diagram of finite type, we define the Artin group associated to $\Gamma$ as in Section \ref{sec:defn_artingroup}. We will then use the group relations presented in this definition to prove the mutation invariance of $A_{\Gamma}$ in Section \ref{sec:main_result}.
